\documentclass[sigplan,10pt,anonymous,nonacm]{acmart}

\settopmatter{printfolios=true,printccs=false,printacmref=false}
\setcopyright{none}
\copyrightyear{2026}
\acmYear{2026}
\acmDOI{}
\acmISBN{}
\acmConference[eBPF '27]{ACM SIGCOMM Workshop on eBPF and Kernel Extensions}{2027}{TBD}

% packages
\usepackage{xspace}
\usepackage{xcolor}
\usepackage{microtype}
\setlength{\emergencystretch}{3em}
\usepackage{array}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,positioning,decorations.pathreplacing}

% commands
\newcommand{\sys}{\textsc{BPFOptBench}\xspace}

\setlength{\textfloatsep}{4pt}
\setlength{\floatsep}{4pt}
\setlength{\intextsep}{4pt}
\setlength{\abovecaptionskip}{2pt}
\setlength{\belowcaptionskip}{-2pt}

\begin{document}

\title{\sys: An Auto-Research Framework and Benchmark for Agentic eBPF Optimization}

\author{Anonymous Authors}
\affiliation{\institution{Anonymous Institution}}
\renewcommand{\shortauthors}{Anonymous}

\begin{abstract}
Kernel extensions such as eBPF are widely deployed on performance-critical
paths for networking, observability, and security. Optimizing these programs
requires cross-layer expertise (verifier semantics, JIT behavior, and workload
characteristics) and offers no reliable local proxy, so the only trustworthy
signal comes from running the full pipeline. This makes eBPF optimization an iterative
search problem where each candidate must be executed to evaluate, a natural fit
for LLM agents. We present \sys, an auto-research framework and benchmark for
eBPF optimization where agents autonomously explore optimization strategies
under real execution feedback. Agents propose actions, the framework executes
real applications under real workloads, and structured observations (verifier
status, JIT output, runtime counters) guide the next iteration. \sys covers 6
production applications (146 programs), exposes action spaces from suite-wide
pass selection to per-program policy, and enforces anti-gaming rules with
hidden correctness checks. Preliminary data motivates autonomous exploration
because static policies yield a near-flat 1.004$\times$ geomean, and local rewrite
counts do not predict speedups, as programs with bytecode changes regress
(1.032$\times$) while unchanged programs improve (0.983$\times$).
\end{abstract}

\keywords{eBPF, agents, optimization, benchmark, verifier, JIT}

\maketitle
\pagestyle{plain}

\section{Background}

\paragraph{eBPF pipeline.}
eBPF programs run on performance-critical paths for networking~\cite{cilium},
security monitoring~\cite{tetragon,falco}, and observability~\cite{bcc}. Each
program passes through a multi-stage pipeline: source code is compiled to BPF
bytecode, loaded into the kernel via the \texttt{bpf()} syscall, verified for
safety by the in-kernel verifier~\cite{linuxkernel-bpf-verifier}, JIT-compiled
to native code, and executed on kernel events. Each stage affects performance
differently, as the verifier may reject semantically correct transformations,
the JIT may expand or contract code depending on architecture, and the workload
determines which paths are hot.

\paragraph{LLM agents for code optimization.}
Recent work demonstrates that LLM agents can leverage execution feedback for
iterative code improvement. KernelBench~\cite{kernelbench2025} evaluates LLMs
on GPU kernel generation and finds that iterative refinement with profiler
feedback improves success rates from 12\% to 43\%.
ComPilot~\cite{compilot2025} uses closed-loop compiler interaction for loop
optimization, achieving 2.66$\times$ speedup on PolyBench.
GSO~\cite{gso2025} benchmarks agents on real-world software optimization,
revealing less than 5\% success rate due to localization challenges. These
works share a common insight: execution feedback is essential because static
analysis and one-shot generation are insufficient.

\section{Motivation}

\paragraph{Why is eBPF optimization hard?}
Optimizing eBPF programs requires cross-layer expertise spanning verifier
semantics, JIT behavior across architectures, and workload characteristics.
Unlike traditional compiler optimization where cost models provide reasonable
proxies, eBPF optimization offers no reliable local signal. Bytecode size does
not predict JIT output because a transformation that reduces bytecode
instructions may expand under JIT due to register allocation or
architecture-specific lowering. JIT size does not predict runtime because
smaller JITed code is not always faster, as inlining a frequently-called helper
increases code size but eliminates call overhead. Microbenchmark gains may not
transfer because an optimization showing 10\% speedup in isolated testing may
have negligible impact on real workloads where the optimized code path is cold.
Architecture dependence further complicates matters because a transformation
profitable on x86-64 may regress on arm64 due to different instruction costs or
branch prediction behavior. The only trustworthy signal comes from running the
full pipeline and measuring actual performance counters.

\paragraph{Existing eBPF optimization.}
Prior work has explored eBPF optimization through different approaches.
K2~\cite{xu2021synthesizing} applies stochastic superoptimization to BPF
bytecode, synthesizing equivalent sequences that may be shorter or faster. K2
demonstrates significant optimization headroom but operates as a one-shot
optimizer without iterative feedback. Merlin~\cite{mao2024merlin} combines LLVM
IR and bytecode transformations, recognizing that eBPF optimization naturally
spans multiple layers, but requires source access and operates at compile time.
The LLVM BPF backend includes static peephole optimizations that combine memory
operations, eliminate redundant moves, and select efficient instruction
encodings, but these are fixed heuristics applied uniformly. These approaches
use fixed search strategies or heuristics, and none evaluates whether an agent
can learn effective optimization strategies from execution feedback, adapting
its policy based on observed outcomes.

\paragraph{Gap.}
Existing agent benchmarks demonstrate the value of execution feedback for code
optimization but do not cover eBPF's unique characteristics, including the
verifier as a hard correctness gate, the JIT as an additional transformation
layer, and production applications with complex workloads. Existing eBPF optimizers use
fixed algorithms rather than adaptive exploration.

\paragraph{Evidence from our corpus.}
Table~\ref{tab:motivation} summarizes preliminary data from our corpus of 6
production applications. Static policies yield near-flat results, and local
rewrite counts do not predict speedups. The most striking finding is the
applied split: programs where bytecode transformations were applied show worse
performance (1.032$\times$ slowdown) than programs left unchanged
(0.983$\times$ speedup). This suggests that the decision of whether to optimize
is as important as how to optimize, motivating an auto-research framework where
agents autonomously explore strategies under real feedback.

\begin{table}[t]
\footnotesize
\centering
\setlength{\tabcolsep}{2pt}
\caption{Corpus data motivates autonomous exploration over static policies.}
\label{tab:motivation}
\begin{tabular}{p{0.15\linewidth}p{0.42\linewidth}p{0.35\linewidth}}
\hline
Evidence & Result & Implication \\
\hline
No-op floors & noop ReJIT 0.902$\times$, skip-ReJIT 0.859$\times$ & noise floor is 0.04$\times$; any claimed improvement must exceed this \\
Static policy & 146 programs, geomean 1.004$\times$, 81 wins / 65 losses & static all-apply policy is near-flat overall \\
Applied split & changed-code programs 1.032$\times$, unchanged 0.983$\times$ & programs where optimization applied actually regress; phase effects dominate \\
Policy tuning & shared subset improved 0.891$\times$ $\to$ 0.898$\times$ & human policy iteration helps but does not solve the corpus \\
\hline
\end{tabular}
\end{table}

\section{Framework Design}

\sys is an auto-research framework enabling agents to iteratively explore eBPF
optimization strategies under real execution feedback. The design goal is to
support autonomous agent exploration while providing the infrastructure for
reproducible benchmarking.

\paragraph{Closed-loop architecture.}
Figure~\ref{fig:loop} shows the \sys closed loop. In each iteration, the agent
observes the current state and prior feedback, then proposes an optimization
action within the allowed action space. The framework executes the action on
real applications under real workloads, the oracle evaluates correctness and
performance, and structured feedback is returned to guide the next iteration.
This architecture enables agents to learn from concrete outcomes rather than
relying on static heuristics, so an agent can observe that a particular pass
causes regressions on networking workloads but improvements on tracing
workloads and adapt its policy accordingly.

\begin{figure}[t]
\centering
\begin{tikzpicture}[
  node distance=0.5cm,
  box/.style={draw, rounded corners=2pt, align=center, minimum width=1.6cm, minimum height=0.5cm, font=\scriptsize},
  arrow/.style={-Latex, thick}
]
\node[box] (agent) {Agent};
\node[box, right=of agent] (action) {Optimization\\Action};
\node[box, right=of action] (exec) {Real Apps\\+ Workloads};
\node[box, below=of exec] (oracle) {Composed\\Oracle};
\node[box, below=of action] (feedback) {Structured\\Feedback};
\draw[arrow] (agent) -- node[above, font=\tiny] {propose} (action);
\draw[arrow] (action) -- node[above, font=\tiny] {execute} (exec);
\draw[arrow] (exec) -- node[right, font=\tiny] {evaluate} (oracle);
\draw[arrow] (oracle) -- node[below, font=\tiny] {score} (feedback);
\draw[arrow] (feedback) -- node[left, font=\tiny] {observe} (agent);
\end{tikzpicture}
\caption{\sys closed loop: agents propose actions, the framework executes on
real applications, and structured feedback guides iteration.}
\Description{Closed loop from agent through action, execution, oracle, feedback.}
\label{fig:loop}
\end{figure}

\paragraph{Optimization layers.}
\sys models eBPF optimization as a hierarchy of artifact layers
(Table~\ref{tab:layers}). The source layer allows restructuring bounds checks
or changing map layouts but requires source access. The LLVM backend layer
enables cost model tuning and peephole patterns but operates at compile time.
The pre-load bytecode layer supports transformations like byte-ladder combining
and static DCE without source access. The live bytecode layer, which is the
primary focus of the first benchmark track, enables pass list selection,
per-program policies, and ReJIT after programs are loaded, leveraging runtime
information such as loaded maps, verifier states, and workload profiles.
Finally, the kernel/JIT layer provides optional backend capabilities like
rotate, conditional select, and bulk memory operations.

\begin{table}[t]
\footnotesize
\centering
\setlength{\tabcolsep}{2pt}
\caption{Optimization layers modeled by \sys. The first benchmark track focuses
on bytecode and live ReJIT layers.}
\label{tab:layers}
\begin{tabular}{p{0.15\linewidth}p{0.45\linewidth}p{0.30\linewidth}}
\hline
Layer & Example actions & First-track role \\
\hline
Source & restructure bounds checks, change map layout, alter tail-call chains & future adapter \\
LLVM backend & BPF cost model tuning, peephole patterns, pass ordering & future adapter \\
Pre-load bytecode & byte-ladder combine, static DCE, helper specialization & supported \\
Live bytecode & pass list selection, per-program policy, ReJIT & primary focus \\
Kernel/JIT & rotate, conditional select, bulk memory, prefetch & optional backend \\
\hline
\end{tabular}
\end{table}

\paragraph{Action space.}
Within each layer, \sys defines action granularity levels ranging from coarse
to fine. At level A0 (no-op), all programs are left unchanged as a baseline. At
level A1 (suite-wide), a fixed pass list is applied to all programs, such as
running \texttt{wide\_mem}, \texttt{map\_inline}, and \texttt{dce} everywhere.
At level A2 (per-app), different policies apply to different applications, for
example enabling \texttt{map\_inline} for Cilium but skipping it for Katran. At
level A3 (per-program), fine-grained control targets individual programs, such
as optimizing only high-run-count programs within an application. Higher levels
A4 (per-site) and A5 (profile-guided) control individual transformation sites
and use PMU data for decisions like branch layout. The first benchmark paper
targets A1--A3 because higher levels require stable per-site reporting and
profile infrastructure.

\section{Benchmark Design}

The benchmark layer defines how agent actions are evaluated. The design goal is
fair, reproducible evaluation with composed correctness and performance checks.

\paragraph{Task model.}
A \sys task specifies five components. The \emph{subject} defines the eBPF
application, workload, target architecture, and optional program subset, using
production applications with representative workloads rather than synthetic
benchmarks to ensure results transfer to real deployments. The \emph{action
space} constrains the allowed optimization layer and granularity, making tasks
tractable while still exercising meaningful decisions. The \emph{budget} limits
maximum attempts, wall-clock time, and token budget, pricing exploration cost
and preventing brute-force search from dominating. The \emph{feedback} returned
after each attempt includes structured observations such as verifier
accept/reject with log excerpts, JIT output size, per-program
\texttt{run\_cnt\_delta} and \texttt{run\_time\_ns\_delta}, and workload-level
metrics, with raw artifacts also available. The \emph{oracle} defines the
composed evaluation criteria described below.

\paragraph{Composed oracle.}
A candidate succeeds only if it passes all oracle gates. The verifier must
accept the transformed program, as rejection means the optimization is invalid
regardless of potential performance. The application must complete its
lifecycle, starting up, attaching all programs, completing the benchmark
workload, and shutting down cleanly without crashes or hangs. Application-specific semantic checks must pass, such as correct packet forwarding for
Cilium, security event detection for Tetragon, and functional load balancing
for Katran. The post-optimization performance must beat the baseline by at
least the noise gate (empirically 0.04$\times$ based on no-op floor
measurements), and the action must not use forbidden shortcuts.

\paragraph{Anti-gaming rules.}
Forbidden actions that would invalidate the benchmark include reducing workload
duration or iteration count, modifying workload inputs or expected outputs,
filtering programs from measurement or ReJIT, hiding or silencing ReJIT
failures, replacing application loaders with direct object loading, and
fabricating or modifying result files. The rate at which agents attempt
measurement invalidation is itself a diagnostic metric.

\paragraph{Metrics.}
\sys reports validity rate (fraction passing verifier and policy checks),
workload success rate (fraction completing workload correctly), performance
success rate (fraction meeting the performance threshold), geomean ratio
(per-program geometric mean of post/baseline runtime), regret (gap to the
best-known valid action), cost (attempts, wall-clock time, tokens consumed),
and integrity failure rate (attempts violating anti-gaming rules).

\section{Preliminary Evaluation}

The evaluation addresses three research questions. RQ1 asks whether eBPF
optimization is a hard closed-loop problem requiring iterative exploration or
whether static policies suffice. RQ2 asks whether \sys correctly exposes
verifier, application, workload, and performance failures as separate outcomes.
RQ3 asks whether closed-loop agents outperform no-op, static, random, and
human-tuned baselines under equal budgets.

\paragraph{Baselines.}
The baselines include no-op (no optimization applied), noop-ReJIT (ReJIT with
identity transformation to control for measurement perturbation), static policy
(applying all available passes uniformly), random search (random pass selection
under the same attempt budget), human-tuned (policy derived from manual corpus
analysis), and oracle best-known (best observed valid action per task, used as
regret reference only).

\paragraph{Current status.}
The \sys framework infrastructure is operational, completing 6 production
applications through native-loader paths: \texttt{bcc/set} (tracing tools),
\texttt{tracee/monitor} (security runtime), \texttt{tetragon/observer}
(security observability), \texttt{katran} (load balancer),
\texttt{cilium/agent} (CNI networking), and \texttt{otelcol-ebpf-profiler}
(continuous profiling). Each run records per-program raw counters, verifier
logs, JIT output sizes, and workload metrics. Table~\ref{tab:motivation}
already demonstrates RQ1 because the near-flat 1.004$\times$ geomean and the
counterintuitive applied-split result show that eBPF optimization requires
adaptive exploration, not fixed heuristics.

\paragraph{Next steps.}
The workshop paper will freeze 20--50 audited tasks with protected workloads
and hidden checks, implement baseline agents including one-shot prompting and
iterative refinement variants, run a scoreboard comparing agent policies under
equal budgets, and analyze failure modes to identify where current agents
struggle.

\bibliographystyle{ACM-Reference-Format}
\bibliography{reference}

\end{document}
